### 2016-07-11 highlights.tex
Add highlights document.

\documentclass{article}
\begin{document}

\section*{Highlights}

\begin{itemize}
\item We develop a framework for analyzing true polar wander (TPW) from a perspective of fluid dynamics.
\item We identify a small set of nondimensional numbers which characterize the rates of true polar wander for a convecting planet.
\item We develop scalings for both the driving forces of TPW and the planetary response in terms of the nondimensional numbers.
\item We present numerical simulations supporting our scaling analyses.
\end{itemize}

\end{document}