### 2016-08-01 frontmatter.tex
Remove abstract from frontmatter.

\chapter{Rates of true polar wander on convecting planets}
\label{chap:tpw_rate}